### 2023-08-14 appendix.tex
up

We relax for now the hypothesis that $\Bad$ is the complement of $\Safe$, thus it is possible for $\Safe$ and $\Bad$ to share closure properties, i.e. for them to be both downward-closed or upward-closed, and we explore both possiblities.

@@

\begin{theorem}\label{down-down}
The {\sc Resilience problem} is undecidable for  effective WSTS with  strong  compatibility such that
%	with more than one minimal element, 
$\Safe=\downarrow \Safe$
and $\Bad=\downarrow \Bad$.
\end{theorem}

@@

\begin{theorem}\label{up-up}
The resilience problem is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
and $\Bad= U \setminus \Safe$ with $U = \uparrow U$.
\end{theorem}

@@

\begin{proposition}
In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the bounded resilience problem is equivalent to the resilience problem.
\end{proposition}

@@

\begin{corollary}\label{B-up-up}
The bounded resilience problem is decidable for WSTS with effective pred-basis,
strong compatibility
 $\Safe=\uparrow \Safe$
and $\Bad=U \setminus \Safe$ with $U = \uparrow U$.
\end{corollary}

\begin{theorem}\label{k-up-up}
The % three %uniform 
$k$-resilience problem is decidable for WSTS with effective pred-basis, strong %upward-
compatibility, $\Safe=\uparrow \Safe$
and $\Bad=U \setminus \Safe$  with $U = \uparrow U$.
%	and \Bad is upward-closed or downward-closed.
\end{theorem}

@@

Most of our undecidability results make use of reset-VASS, which are WSTS, and for which several problems are undecidable, including reachability~\cite{araki1976PN, dufourd1998reset}. This has consequences for {\sc Resilience}. For instance in the case $\Safe = \downarrow \Safe$, it is possible for $\Safe$ to be a singleton element. Even then, though, {\sc Resilience} (resp. {\sc State-resilience}) is not strictly reachability, and we will need to delve more into the details to obtain undecidability results.

@@

\subsection{Home-state is undecidable for Minsky machine, reset-VASS and thus, WSTS}\label{HS-Minsk}

We mentionned in Section~$3$ that {\sc Resilience} is undecidable for WSTS in general when $\Safe = \downarrow \Safe$ hence $\Bad = \uparrow \Bad$. Let us provide a more detailed undecidability proof here.

@@

\problemx{Home-state}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and a state $s \in S$.}
{$S \longrightarrow^{*} \{s\}$ ?\newline}

% whether a state $s$ is such that, for all state $s_0$ there exists a path from $s_0$ to $s$.

Home-state is undecidable for Minsky machine with $3+$ counters.
This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61, Min67}.
From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
such that the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter. We build $M'$ to simulate $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then finally lower the third counter until it reaches $0$.
% Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
% or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
Thus {\sc Home-state} is undecidable for Minsky machines, and hence {\sc Resilience} is undecidable.

@@

We are now interested in the 
%following
 undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in Reset $d$-VASS, which consists in, given a
Reset $d$-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
%
We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live, that is, whether it is always ,
which we then reduce to {\sc State-resilience}.
A control-transition $t$ of a Reset $d$-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
 $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole Reset $d$-VASS is live if all its control-transitions are
live. This leads to the following problem.

@@

\section{More Appendix thing if necessary}\label{appendix W}

>>>>

We relax now the hypothesis that $\Bad$ is the complement of $\Safe$. In this appendix it is thus possible for $\Safe$ and $\Bad$ to share closure properties, i.e. for them to be both downward-closed or upward-closed. We will now explore both possiblities.

@@

\begin{theorem}\label{down-down}
{\sc Resilience} is undecidable for  effective WSTS with  strong  compatibility such that
%	with more than one minimal element, 
$\Safe=\downarrow \Safe$
and $\Bad=\downarrow \Bad$.
\end{theorem}

@@

\begin{theorem}\label{up-up}
{\sc Resilience} is decidable for effective WSTS with effective pred-basis, $\Safe=\uparrow \Safe$
and $\Bad= U \setminus \Safe$ with $U = \uparrow U$.
\end{theorem}

@@

\begin{proposition}
In WSTS with strong compatibility and effective pred-basis,  $\Safe=\uparrow \Safe$, the {\sc Bounded resilience} is equivalent to {\sc Resilience}.
\end{proposition}

@@

\begin{corollary}\label{B-up-up}
{\sc Bounded resilience}  is decidable for WSTS with effective pred-basis,
strong compatibility
 $\Safe=\uparrow \Safe$
and $\Bad=U \setminus \Safe$ with $U = \uparrow U$.
\end{corollary}

\begin{theorem}\label{k-up-up}
{\sc $k$-resilience}  is decidable for WSTS with effective pred-basis, strong %upward-
compatibility, $\Safe=\uparrow \Safe$
and $\Bad=U \setminus \Safe$  with $U = \uparrow U$.
%	and \Bad is upward-closed or downward-closed.
\end{theorem}

@@

Most of our undecidability results rely on reset-VASS, which are WSTS, and for which several problems are undecidable, including reachability~\cite{araki1976PN, dufourd1998reset}. This has consequences for {\sc Resilience}; for instance in the case $\Safe = \downarrow \Safe$, it is possible for $\Safe$ to be a singleton element. Even then, {\sc Resilience} (resp. {\sc State-resilience}) is not strictly speaking the same as reachability, and we will need to delve more into the details to obtain undecidability results.

@@

\subsection{Home-state is undecidable for Minsky machine, reset-VASS and WSTS}\label{HS-Minsk}

We mentionned in Section~$3$ that {\sc Resilience} is undecidable for WSTS in general when $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$. Let us provide a more detailed undecidability proof here.

@@

\problemx{Home-state}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and a state $s \in S$.}
{$\post^*(S) \subseteq \pred^*(\{s\})$ ?\newline}

% whether a state $s$ is such that, for all state $s_0$ there exists a path from $s_0$ to $s$.
% post ∗ (S ) ⊆ pred ∗ (s_0)

Home-state is undecidable for Minsky machine with $3+$ counters.
This stems from the fact $2$-counter Minsky machine termination is undecidable~\cite{Min61, Min67}.
From a $2$-counter Minsky machine $M$, one can construct a $3$-counter Minsky machine $M'$ 
such that the Minsky machine $M$ terminates for all inputs iff the Minsky machine $M'$ can reach $(0,0,0)$ from any input with at least $1$ on its third counter. We build $M'$ to simulate $M$ until it reaches a control state indicative of termination, then lower the first two counters until they reach $0$, then finally lower the third counter until it reaches $0$.
% Home-state indécidable pour les Minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
% or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.
Thus {\sc Home-state} is undecidable for Minsky machines. Since $\post^*(S) \subseteq \pred^*(\{s\})$ is equivalent to $S \rightarrow^* \{s\}$,  {\sc Home-state} is a particular instance of {\sc Resilience}. Hence {\sc Resilience} is undecidable for Minsky machines.

@@

We are now interested in the 
%following
 undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in Reset VASS, which consists in, given a
Reset VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
%
We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live, that is, whether it is always ,
which we then reduce to {\sc State-resilience}.
A control-transition $t$ of a Reset $d$-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
 $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole Reset $d$-VASS is live if all its control-transitions are
live. This leads to the following problem.

@@
